A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-10-10 paper/paper.tex
paper: macros; minor polish

% Universe, referred to as U
\newcommand{\U}[1]{\ensuremath{\mathsf{\color{spanish-blue}{HSet}}_{#1}}}

@@

% Some abbreviations
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}

@@

\subsection{Homotopy type theory}
Homotopy type theory (HoTT) is a logical foundation based on Martin-Löf type theory and the univalence axiom~\cite{hottbook}. HoTT is based on the notion of a basic \textsf{Type}, and the notion of sets can be recovered as \textsf{HSet}, where \textsf{HSet} is a straightforward packaging of two fields:

\begin{align*}
  \textsf{Dom} & : \textsf{Type}                                          \\
  \textsf{UIP} & : \forall x y : \textsf{Dom}, \forall p q : x = y, p = q
\end{align*}

In the \textsf{HSet} universe, three properties hold:

@@

\textsf{HSet} provides us a form of unicity of identity of proofs (UIP), which is a form of proof irrelevance:

@@

The latter correspondence can be expressed in HoTT, for $T: \textsf{Type}$, as an equivalence between $\Sigma S: \textsf{Type}\;.\;S \rightarrow T$ and $T \rightarrow \textsf{Type}$, which can be proved using univalence. This correspondence also holds on \textsf{HSet}, for $T: \textsf{HSet}$:

\begin{equation*}
  \Sigma S: \textsf{HSet}\;.\;S\rightarrow T \simeq T \rightarrow \textsf{HSet}
\end{equation*}

Let us now come back to cubical sets. A cubical set can be represented as a sequence of fibrations, together with appropriate coherence conditions:

\begin{equation*}
  \begin{tikzcd}
    X_0 : \U{} & X_1 : \U{} \arrow[l, "\delta^\epsilon_0" description, shift left=2] \arrow[l, "\delta^\omega_0" description, shift right=2] & X_2 : \U{} \arrow[l, "\delta^\epsilon_1" description, shift left=6] \arrow[l, "\delta^\epsilon_0" description, shift left=2] \arrow[l, "\delta^\omega_0" description, shift right=2] \arrow[l, "\delta^\omega_1" description, shift right=6] & \ldots
  \end{tikzcd}
\end{equation*}

We can iteratively apply the above correspondence to obtain an indexed representation of cubical sets:

\begin{align*}
  X_0 & : \U{}                                                                                                              \\
  X_1 & : X_0 \times X_0 \rightarrow \U{}                                                                                   \\
  X_2 & : \forall a b c d, X_1 (a, b) \rightarrow X_1 (a, c) \rightarrow X_1 (b, d) \rightarrow X_1 (c, d) \rightarrow \U{} \\
  \ldots
\end{align*}

@@

The proof of $\cohf$ require a higher-dimensional coherence condition which we obtain by working in \texttt{HSet}. If the proofs of the same equality were not equated, there would be a need for arbitrary many higher-dimensional coherences (see e.g.~\cite{herbelin15} for a discussion on the de facto need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory).

>>>>

% Universe, referred to as U
\newcommand{\U}[1][]{\ensuremath{\mathsf{\color{spanish-blue}{HSet}}_{#1}}}

@@

% Some abbreviations
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommand{\Type}{\textsf{Type}}
\newcommand{\Dom}{\textsf{Dom}}
\newcommand{\UIP}{\textsf{UIP}}

@@

\subsection{Homotopy type theory}
Homotopy type theory (HoTT) is a logical foundation based on Martin-Löf type theory and the univalence axiom~\cite{hottbook}. HoTT is based on the notion of a basic Type, and the notion of sets can be recovered as \U, where $\U$ is a straightforward packaging of two fields:

\begin{align*}
  \Dom & : \Type                                          \\
  \UIP & : \forall x y : \Dom, \forall p q : x = y, p = q
\end{align*}

In the $\U$ universe, three properties hold:

@@

$\U$ provides us a form of unicity of identity of proofs (UIP), which is a form of proof irrelevance:

@@

The latter correspondence can be expressed in HoTT, for $T: \Type$, as an equivalence between $\Sigma S: \Type\;.\;S \rightarrow T$ and $T \rightarrow \Type$, which can be proved using univalence. This correspondence also holds on \U, for $T: \U$:

\begin{equation*}
  \Sigma S: \U\;.\;S\rightarrow T \simeq T \rightarrow \U
\end{equation*}

Let us now come back to cubical sets. A cubical set can be represented as a sequence of fibrations, together with appropriate coherence conditions.

\begin{equation*}
  \begin{tikzcd}
    X_0 : \U & X_1 : \U \arrow[l, "\delta^\epsilon_0" description, shift left=2] \arrow[l, "\delta^\omega_0" description, shift right=2] & X_2 : \U \arrow[l, "\delta^\epsilon_1" description, shift left=6] \arrow[l, "\delta^\epsilon_0" description, shift left=2] \arrow[l, "\delta^\omega_0" description, shift right=2] \arrow[l, "\delta^\omega_1" description, shift right=6] & \ldots
  \end{tikzcd}
\end{equation*}

We can iteratively apply the above correspondence to obtain an indexed representation of cubical sets.

\begin{align*}
  X_0 & : \U                                                                                                              \\
  X_1 & : X_0 \times X_0 \rightarrow \U                                                                                   \\
  X_2 & : \forall a b c d, X_1 (a, b) \rightarrow X_1 (a, c) \rightarrow X_1 (b, d) \rightarrow X_1 (c, d) \rightarrow \U \\
  \ldots
\end{align*}

@@

The proof of $\cohf$ require a higher-dimensional coherence condition which we obtain by working in \U. If the proofs of the same equality were not equated, there would be a need for arbitrary many higher-dimensional coherences (see e.g.~\cite{herbelin15} for a discussion on the de facto need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory).
